\chapter{Introduction}

Following the massive expansion of embedded systems led by smart devices, a new inevitable trend began to develop. The automotive industry started to embed more and more computational power in the ubiquitous personal car. Automotive specific Linux operating systems  became unavoidable. For this particular reason a few projects started to emerge that aimed to shape the future development efforts towards a maintainable, safe and uniform ecosystem. They are initiatives started by colossal companies whose main line of activity is not related to software but understood the importance of a high quality software component.

While automotive giants discussed how to better approach a future of connected cars, the Linux Foundation came with a proposition - a specification for Automotive Grade Linux (AGL). AGL is intended to be a distribution that serves as a standard of quality. It follows the model of an earlier, very successful specification, the Carrier Grade Linux.

In a context of ever expanding embedded computer networks and the emergence of new specifications and standards, the Linux ecosystem is still lacking open source tools that facilitate testing. While a lot of effort is directed towards specialized and efficient testing tools, there is currently no solution that offers a full image about the compliance of a distribution with a specification. This paper describes a framework proposal for this particular task - validate a whole Linux distribution based on a given specification. The framework aims to contribute to an efficient, well thought and safe embedded Linux ecosystem by reducing the development effort through automation. It anticipates an increase of Linux distribution or application vendors and tries to enable a community-driven effort towards higher standards that will ultimately benefit the end-users.

For convenience, the framework proposal will also be referred to as "Ellida", the name chosen for the software implementation\footnote{In the old Norse culture, Ellida is the name of a magical ship offered to the Vikings by the gods - the first Viking dragon ship.}

